Mergeable Types

### 2017-02-28 paper/intro.tex
edits

In this paper, we consider how declarative abstractions can be used to
overcome this tension to provide the best of both worlds: functional
programs whose logical structure can be seamlessly transplanted to a
distributed setting, while nonetheless being resilient to system
realities such as network failure and latency.  Our key insight is the
development of a principled methodology that allows us transform
low-level operational details such as replication and consistency
into high-level compositional reasoning on language-level datatypes.

@@

Our main innovation is thus the development of a programming model
that realizes a monadic version control system centered around data,
rather than file, coherence.  The \name monad lets programmers write
and compose concurrent computations around multiple (implicit)
versions of a mergeable datatype, an ordinary ML datatype additionally
equipped with a \emph{merge} function responsible for deriving a
consistent global state from a collection of local versions of that
state.  A computation progresses by forking (i.e., replicating) new
versions of existing versions, allowing synchronization-free local
computation to proceed on those nodes, creating new versions along an
existing \emph{branch} that represents updated local instances, and by
merging branches to realize global consistency.

>>>>

In this paper, we consider how declarative abstractions can be used to
overcome this tension to provide the best of both worlds: functional
programs whose logical structure can be seamlessly transplanted to a
distributed setting, while nonetheless being resilient to system
realities such as network failure and latency.  Our key insight is the
development of a principled methodology that allows us to transform
low-level operational details such as replication and consistency into
high-level compositional abstractions over language-level datatypes.

@@

Our main innovation is thus the development of a programming model
that realizes a monadic version control system centered around data,
rather than file, coherence.  The \name (Versioned ML) monad lets
programmers write and compose concurrent computations around multiple
(implicit) versions of a mergeable datatype, an ordinary ML datatype
additionally equipped with a \emph{merge} function responsible for
deriving a consistent global state from a collection of local versions
of that state.  A computation progresses by forking (i.e.,
replicating) new versions of existing versions, allowing
synchronization-free local computation to proceed on those nodes,
creating new versions along an existing \emph{branch} that represents
updated local instances, and by merging branches to realize global
consistency.
